---
title: "Time Series with dygraphs"
output:
  html_document:
    fig_width: 6
    fig_height: 4
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, fig.width=6, fig.height=2.5}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures", ylab = "Temp (F)") 
```


## Simple demos

```{r}
library(dygraphs)
lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths)
```


Note that this graph is fully interactive: as your mouse moves over the series individual values are displayed. You can also select regions of the graph to zoom into (double-click zooms out).

You can customize dygraphs by piping additional commands onto the original dygraph object. Here we pipe a dyRangeSelector onto our original graph:

```{r}
dygraph(lungDeaths) %>% dyRangeSelector()
```


Note that this example uses the %>% (or “pipe”) operator from the magrittr package to compose the dygraph with the range selector. You use a similar syntax to customize axes, series, and other options. For example:

```{r}
dygraph(lungDeaths) %>%
  dySeries("mdeaths", label = "Male") %>%
  dySeries("fdeaths", label = "Female") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height = 20)
```

Many options for customizing series and axis display are available. It’s even possible to combine multiple lower/value/upper style series into a single display with shaded bars. Here’s an example that illustrates shaded bars, specifying a plot title, suppressing the drawing of the grid for the x axis, and the use of a custom palette for series colors:


```{r}
hw <- HoltWinters(ldeaths)
predicted <- predict(hw, n.ahead = 72, prediction.interval = TRUE)

dygraph(predicted, main = "Predicted Lung Deaths (UK)") %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dySeries(c("lwr", "fit", "upr"), label = "Deaths") %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Set1"))
```

Various features are available [here](http://rstudio.github.io/dygraphs/index.html)