---
title: "Polynomial Model"
author: "Benoit Liquet"
date: "November 1, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Polynomial regression

In a polynomial model, the relationship between the explained variable $Y$ and an explanatory variable $X$ is represented in a non-linear fashion, such as:
\[Y=\beta_0+\beta_1X+\beta_2X^2+\ldots+\beta_pX^p+\epsilon.\]
This model is a multiple regression model with $p$ regressors: the powers of the explanatory variable.

To perform regression in a polynomial model, you simply need to correctly specify the formula associated with the model in the  function **lm()**. Two R functions are useful: **I()** and  **poly()**. The next table gives a few examples of formulae for polynomial models.


```{r, results = "asis", echo = FALSE, message = FALSE}
library(knitr)

tex2markdown <- function(texstring) {
  writeLines(text = texstring,
             con = myfile <- tempfile(fileext = ".tex"))
  texfile <- pandoc(input = myfile, format = "html")
  cat(readLines(texfile), sep = "\n")
  unlink(c(myfile, texfile))
}

textable <- "
\\begin{table}
\\begin{tabular}{ll}
\\hline
\\textbf{Model}&  \\textbf{R formula} \\\\
\\hline\\hline
$M_1$: $Y=\\beta_0+\\beta_1X+\\beta_2X^2$ & Y$\\sim$poly(X,2,raw=TRUE)\\\\
$M_2$: $Y=\\beta_1X+\\beta_2X^2+\\beta_3X^3$ & Y$\\sim$-1+poly(X,3,raw=TRUE)\\\\
$M_3$: $Y=\\beta_0+\\beta_1X+\\beta_2X^3$ & Y$\\sim$\\verb#X+I(X^3)#\\\\
$M_4$: $Y=\\beta_1X+\\beta_2X^3+\\beta_3X^4$ & Y$\\sim$\\verb#-1+X+I(X^3)+ I(X^4)#\\\\
\\hline
\\end{tabular}
\\end{table}
"

tex2markdown(textable)
```



## Example on synthetic data

Suppose you are asked to propose a model to predict a  variable $Y$ given an explanatory variable $X$. You are given a sample of size~$n$. 



- Download the file http://biostatisticien.eu/springeR/fitpoly.RData in your current directory.

```{r}
load("fitpoly.RData")
attach(fitpoly)
```

- Draw a scatter plot of variable $Y$ as a function of variable $X$.

```{r}
plot(Y~X)
```

- Is there a linear relationship between these two variables? Fit a regression line on the previous plot.

```{r}
lin.model <- lm(Y~X)
summary(lin.model)
plot(Y~X)
abline(lin.model)
```

- Perform polynomial regression to fit the data better.

```{r}
poly.model <- lm(Y~poly(X,3,raw=TRUE),data=fitpoly)
summary(poly.model)
```

- Draw the estimated polynomial over the scatter plot. Draw the confidence curve for the mean of $Y$ for $X\in[-3.5,3.5]$. Add the prediction interval of the model for $X\in[-3.5,3.5]$.

```{r}
coef <- coef(poly.model)
plot(Y~X)
curve(coef[1]+coef[2]*x+coef[3]*x^2+coef[4]*x^3,add=TRUE)
x <- seq(-3.5,3.5,length=50)
pred.int <- predict(poly.model,data.frame(X=x),
                    interval="prediction")[,c("lwr","upr")]
conf.int <- predict(poly.model,data.frame(X=x),
                    interval="confidence")[,c("lwr","upr")]
matlines(x,cbind(conf.int,pred.int),lty=c(2,2,3,3),
         col=c("red","red","blue","blue"),lwd=c(2,2,1,1))
legend("bottomright",lty=c(2,3),lwd=c(2,1),
       c("confidence","prediction"),col=c("red","blue"))
```